---
title: "Research Questions"
#bibliography: references.bib
author: "Paige Kemper"
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()


colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }


```


```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```



------------------------------------------------------------------------
`r colorize("Research Questions Tab", "red")`

------------------------------------------------------------------------

### **Preliminary Research Questions**
What impact does ethnic diversity in departments have on cross department collaboration?

Do the researcher’s # of collaborations within department vs. across departments differ depending on their ethnic diversity across Dutch universities? 

Do the department's proportion of within-department collaboration vs beyond department collaboration change depening on the ethnic diversity of the department across universities in the netherlands? 

How is this impacted by how long researchers have been at the university? 



------------------------------------------------------------------------
`r colorize("Revised Research Questions", "red")`

------------------------------------------------------------------------

### **Updated Research Questions**

-   What impact does ethnic diversity within in departments have on cross department collaboration? –\> **How does ethnic/nationality diversity within academic departments influence collaboration patterns among researchers in dutch universities, and how is this relationship moderated by researchers' institutional tenure?**


notes from class: 

-   _ethnic diversity doable manually -- but macro level -- department level -- need to include a lot of departments to compare _
 
-   _descriptive question for all departments, and then pick 2 departments (high in diversity vs low in diversity (intersectionality)) to compare -- think of creative ways to use the data we have, ex. native dutch vs non-native dutch_
 
-   _affiliations observed in open alex -- if ethnicity of scholar is difficult, look at if first affiliation is non-dutch university!!! as a proxy_ 


-- list describptive questions: clusters, clustering by dept? clust
 

Additional research questions/sub-research questions:
 
    -   How does ethnic diversity impact individual researchers proportion of inter-department collaborations as opposed to cross department (or cross-univerisyt) collaborations in Dutch universities?

    -   How does inter-department ethnic diversity impact intra-department collaborations, across universities in the Netherlands?

    -   How are these effects amplified or reduced by the time the researcher has been at the department (or mean tenure of department)?


```{r}
## ref
```

